---
title: "World cup goals"
author: "Mine Çetinkaya-Rundel"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-pkg}
library(tidyverse)
```

The FIFA World Cup is a global football competition contested by the various 
football-playing nations of the world. It is contested every four years, match 
data from each game is maintained on the FIFA World Cup Archive website.

```{r load-data}
wc <- read_csv("data/WorldCupMatches-01.csv")
```

The figure below shows the total number of goals scored during world cup games 
between `r min(wc$year)` and `r max(wc$year)`. We can see that the total number 
of goals have been increasing over the year.

```{r calc-totals}
wc_tots <- wc %>%
  mutate(total_goals = home_team_goals + away_team_goals) %>%
  group_by(year) %>%
  summarise(total_goals_year = sum(total_goals))
```

```{r plot-totals}
ggplot(data = wc_tots, aes(x = year, y = total_goals_year)) +
    geom_point() +
    geom_line() +
    geom_smooth() +
    labs(
      x = "Year", y = "Total goals",
      title = "Total goals during each World Cup"
    )
```
