---
title: "Getting started"
subtitle: "Teaching Data Science, Reproducibly <br> @ ICOTS 2018"
author: "Mine Çetinkaya-Rundel & Colin Rundel"
date: "July 7, 2018"
output:
  xaringan::moon_reader:
    css: "../slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

class: center, middle

# Getting started

---

class: middle

.question[
What do you do on the first day of class? What do you not do?
]

---

## Cherish day one!

.pull-left[
**Don't do this:**
- Install R
- Install RStudio
- Install the following packages:
    - rmarkdown
    - tidyverse
    - …
- Load these packages
- Install git
]
.pull-right[
**Do this:**
- Go to [rstudio.cloud](https://rstudio.cloud/) (or some other server based solution)
- Log in with your ID & pass  

`> hello R!`
]

---

## Don't start with this

```{r lame-start}
class(mtcars$mpg)
mean(mtcars$mpg)
median(mtcars$mpg)
sd(mtcars$mpg)
```

---

## Start with (something like) this

```{r better-start, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5.5, dpi=150}
library(tidyverse)
library(unvotes)
library(lubridate)
un_votes %>%
  filter(country %in% c("United States of America", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```

---

## Let's do it!

- **Go to** http://bit.ly/teach-data-sci-rscloud and join the workspace
- **Make a copy** of the project called *UN votes*
- **Open** the file called `unvotes.Rmd` and **click** on Knit HTML
- **Discuss** with the person sitting next to you what you see in the two figures. 
Don't worry about the code to produce them just yet, only focus on the output.  

--
- In the last figure, **change** `"Turkey"` to a country of interest to you and 
recreate the figure by knitting the document again. Do this by making a change 
in Line 95 of the document only.
    - Tip: The country you choose must be in the dataset and must be spelled 
    exactly the same way as it appears in the dataset. You can use the 
    searchable table in the Appendix to locate the country of your choice and 
    to confirm its spelling.
- **Discuss** with the person sitting next to you why you chose this country and 
how its voting record is similar to or different than the US.
